% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={NeuroBeat: Energy-Accounted Spiking Networks for Honest Inter-Patient Ventricular Ectopic Beat Detection},
  pdfauthor={Valerio Merlini (independent researcher)},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{NeuroBeat: Energy-Accounted Spiking Networks for Honest
Inter-Patient Ventricular Ectopic Beat Detection}
\author{Valerio Merlini (independent researcher)}
\date{July 2026}

\begin{document}
\maketitle

\emph{Contact:} github.com/Talch87/neuro-beat ~·~ \emph{Code, weights,
live dashboard:} talch87.github.io/neuro-beat

\emph{Status: preprint draft, content-complete. All results are from
locked 5-seed runs (val-locked, DS2 frozen). Pending before posting: a
final author-list check on the SNN-ECG references marked †.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Abstract}\label{abstract}

Spiking neural networks (SNNs) are attractive for always-on cardiac
monitoring because their event-driven computation maps onto low-power
neuromorphic hardware. Yet much of the SNN-ECG literature reports
accuracy under protocols that quietly inflate it --- intra-patient
splits, or decision thresholds tuned on the test set --- and rarely
accounts for the energy that motivates the approach in the first place.
We present \textbf{NeuroBeat}, a compact leaky-integrate-and-fire (LIF)
SNN for ventricular ectopic beat (VEB) detection, evaluated under a
deliberately conservative protocol: a patient-disjoint inter-patient
split (MIT-BIH DS1→DS2), an operating point fit \textbf{only} on a DS1
validation holdout and then \textbf{frozen} for all test data, and an
explicit per-beat synaptic-operation (SynOps) energy budget. At a single
operating point the model faces a genuine sensitivity--precision
tradeoff (VEB sensitivity 0.89 at PPV 0.49, \emph{or} 0.86 at 0.68 ---
not both), which we trace to calibration variance rather than model
capacity. We break it with a \textbf{gated-ensemble two-stage cascade}:
a genuinely sparse high-recall screener (5,987 SynOps/beat) runs on
every beat and gates a 3-seed ensemble confirmer that runs only on the
\textasciitilde27\% flagged candidates. The cascade reaches \textbf{VEB
sensitivity 0.923 at PPV 0.616 on DS2 within 23,385 SynOps/beat} ---
meeting sensitivity ≥0.90, PPV ≥0.60, and the ≤25k energy budget
simultaneously --- and holds VEB sensitivity ≥0.90 across all three
databases (DS2, SVDB, INCART) under one frozen operating point. Along
the way we show that count-pooled encoding makes energy
timestep-independent, so time resolution is a free discrimination lever
but short networks are not cheap. Finally, we report a negative result:
supraventricular (SVEB) detection is data- and lead-limited on a single
lead (DS2 SVEB sensitivity \textasciitilde0.13, rising to 0.62 on
12-lead INCART), which we address separately with an SVEB specialist
rather than by compromising the VEB operating point. Code, weights, and
a live results dashboard are public.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Introduction}\label{introduction}

Continuous ECG monitoring on wearable and implantable devices is
power-limited: the detector must run for days on a coin cell.
Ventricular ectopic beats (VEBs) are clinically important --- frequent
VEBs and ventricular runs precede dangerous arrhythmias --- and a good
always-on VEB detector must be both \textbf{sensitive} (miss few true
VEBs) and \textbf{cheap} (little energy per beat).

Spiking neural networks are a natural fit: they compute with sparse
binary events, and on neuromorphic substrates energy scales with the
number of synaptic operations actually performed, not with a fixed
clock. This has driven a wave of SNN-ECG work. However, two problems
recur:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Optimistic evaluation.} Reported accuracy is often inflated by
  intra-patient splits (the same patient's beats appear in train and
  test) or by selecting the decision threshold on the test set. Both
  leak information that a deployed device never has.
\item
  \textbf{Missing energy accounting.} The energy argument is the reason
  to use an SNN, yet many papers report only accuracy, or parameter
  counts, with no per-inference operation budget.
\end{enumerate}

We take the opposite stance on both. Our contributions:

\begin{itemize}
\tightlist
\item
  \textbf{An honest, val-locked inter-patient protocol.} We use the de
  Chazal DS1/DS2 patient-disjoint split, fit the operating point
  (per-class logit biases) on a DS1 validation holdout, and freeze it
  for DS2 and all external data. The test set chooses nothing.
\item
  \textbf{Explicit energy accounting.} We report per-beat SynOps and
  hold every configuration to a ≤25k-SynOps/beat budget. We identify
  \textbf{time resolution} (number of SNN timesteps) as a cheap
  discrimination lever.
\item
  \textbf{Cross-database external validation.} The single frozen
  operating point is applied unchanged to two databases the model never
  trained on (MIT-BIH SVDB, INCART), quantifying real generalization.
\item
  \textbf{A two-stage energy-gated cascade} that breaks the
  single-operating-point sensitivity--precision tradeoff at low average
  energy.
\item
  \textbf{A characterized negative result} for supraventricular beats:
  SVEB is data-/lead-limited, not an architecture failure, and is better
  pursued by a dedicated specialist than by degrading the VEB operating
  point.
\end{itemize}

Everything is reproducible: code, frozen weights, per-seed logs, and a
public dashboard.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Related Work}\label{related-work}

\textbf{Inter-patient benchmarks and evaluation hygiene.} The AAMI EC57
standard {[}AAMI-EC57{]} defines the five beat classes and the
sensitivity/PPV reporting convention. de Chazal et
al.~{[}deChazal2004{]} introduced the patient-disjoint DS1/DS2 split (44
MIT-BIH records, \textasciitilde50k beats each) that we adopt. A 2025
systematic review {[}Silva2025{]} surveys 2017--2024 arrhythmia
classifiers and finds that, despite high reported accuracy, few studies
implement patient-independent partitioning or account for embedded
constraints --- the exact gap our protocol targets. NeuroBeat
contributes an explicitly val-locked operating point (fit on a DS1
holdout, never on DS2) and cross-database external validation on top of
the DS1/DS2 convention.

\textbf{SNNs for ECG / arrhythmia.} SNNs are attractive for ECG because
event-driven computation maps onto low-power neuromorphic hardware.
Prior work reports strong MIT-BIH accuracy with delta-modulation
encoding of the ECG and its derivatives {[}SNN-ECG-BSPC2021{]},
spike-driven neuromorphic processors for wearable ECG {[}Chu2022{]},
hardware/software co-design {[}SparrowSNN2024{]}, and axonal-delay SNNs
{[}AxonalDelays2025{]}. Our encoding (count-pooled delta channels over
signal orders) is in this family, but most of these report intra-patient
or test-tuned accuracy; we instead report val-locked inter-patient
numbers with explicit energy budgets and a negative result for the
supraventricular class.

\textbf{SNN training and energy.} We train with surrogate-gradient
backpropagation-through-time {[}Neftci2019{]} via snnTorch
{[}Eshraghian2023{]}. On neuromorphic substrates, per-inference energy
scales with synaptic operations; reported ECG figures show
\textasciitilde4.68 µJ/inference on a neuromorphic chip versus
\textasciitilde450 µJ on a conventional processor {[}Chu2022{]}, and
general-purpose neuromorphic cores such as Loihi {[}Davies2018{]}
motivate the SynOps proxy we budget against.

\textbf{Cascaded / gated inference.} Two-stage screener--confirmer and
energy-gated designs are standard elsewhere for trading average cost
against accuracy; we apply the pattern to keep an ensemble-grade
confirmer affordable by running it only on flagged candidates.

\textbf{Data.} MIT-BIH {[}Moody2001{]}, its Supraventricular database
(SVDB), and the St.~Petersburg INCART 12-lead database are distributed
via PhysioNet {[}Goldberger2000{]}.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Data}\label{data}

\textbf{MIT-BIH Arrhythmia Database (primary).} 48 two-lead records, 360
Hz. Beats are mapped to the five AAMI classes: N (normal), S
(supraventricular ectopic, ``SVEB''), V (ventricular ectopic, ``VEB''),
F (fusion), Q (unknown). We use the de Chazal inter-patient split:
\textbf{DS1} for training, \textbf{DS2} for test, with no patient in
both. From DS1 we hold out records \textbf{201, 207, 223} as a
patient-disjoint \textbf{validation} set for operating-point selection.

\textbf{External databases (test-only generalization).} MIT-BIH
Supraventricular Arrhythmia Database (\textbf{SVDB}, 128 Hz) and
St.~Petersburg INCART 12-lead database (\textbf{INCART}, 257 Hz).
Signals are resampled to 360 Hz (polyphase,
\texttt{scipy.signal.resample\_poly} on the reduced gcd ratio) and
annotation sample positions rescaled accordingly. Lead selection:
MIT-BIH and SVDB lead 0; INCART lead II (index 1).

\textbf{Beat segmentation and RR features.} Beats are windowed around
each annotated R-peak. For each beat we also compute three
patient-normalized RR-interval features (previous, current, and ratio),
standardized using training-set statistics only.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Methods}\label{methods}

\subsubsection{4.1 Spike encoding}\label{spike-encoding}

Each beat window is converted to a spike tensor by a
\textbf{count-pooled delta encoding}. For each signal order in
\texttt{orders} (order 0 = the signal, order 1 = its first derivative),
threshold crossings of magnitude \texttt{θ\ =\ 0.12} are pooled into
\texttt{T} time bins, producing two channels (up/down) per order. With
\texttt{orders\ =\ {[}0,\ 1{]}} this yields
\texttt{2·\textbar{}orders\textbar{}\ =\ 4} input channels over
\texttt{T} timesteps. Time resolution \texttt{T} is a key lever: raising
it sharpens discrimination at almost no extra energy (Section 6).

\subsubsection{4.2 Network}\label{network}

A compact two-layer LIF network (\texttt{SNNClassifier}): \texttt{fc1}
projects the 4 input channels to \texttt{H\ =\ 128} hidden LIF units;
\texttt{fc2} projects hidden spikes to the 5 class logits. The three RR
features are projected once through a dense map into the hidden state.
Readout is by accumulated membrane potential. The network is trained by
surrogate-gradient backpropagation-through-time (snnTorch) with Adam (lr
4e-3, 100 epochs) and \texttt{sqrt}-scaled class weights to counter
class imbalance.

\subsubsection{4.3 Val-locked operating
point}\label{val-locked-operating-point}

The network outputs 5 logits; the decision applies a fixed per-class
\textbf{bias} vector \texttt{b\ =\ {[}0,\ b\_S,\ b\_V,\ −12,\ −12{]}} (F
and Q suppressed) and takes the argmax. \texttt{(b\_S,\ b\_V)} is chosen
on the \textbf{DS1 validation holdout only} and then \textbf{frozen} for
DS2 and every external database. We consider three val-fit strategies:

\begin{itemize}
\tightlist
\item
  \textbf{sens-first} --- maximize VEB PPV subject to VEB sensitivity ≥
  0.90;
\item
  \textbf{ppv-first} --- maximize VEB sensitivity subject to VEB PPV ≥
  0.60;
\item
  \textbf{balanced} --- maximize VEB F1.
\end{itemize}

The frozen NeuroBeat-VEB v1 uses \textbf{sens-first} (missing
ventricular beats is the costlier error), and we report the full
sensitivity--PPV frontier so the tradeoff is explicit.

\subsubsection{4.4 Energy proxy}\label{energy-proxy}

Per-beat energy is the synaptic-operation (SynOps) count:
\texttt{SynOps\ =\ (Σ\_t\ input\_spikes)·H\ +\ (Σ\_t\ hidden\_spikes)·C\ +\ n\_RR·H},
where \texttt{C\ =\ 5} classes and the last term is the one-shot
RR→hidden projection. The first term dominates and is
encoding-dependent, which is why sparse, low-order input and modest
\texttt{T} matter. Budget: \textbf{≤ 25,000 SynOps/beat}.

\subsubsection{4.5 Two-stage cascade}\label{two-stage-cascade}

To obtain high sensitivity \textbf{and} high precision without paying
full energy on every beat, we cascade two independent networks:

\begin{itemize}
\tightlist
\item
  \textbf{Stage 1 (screener)} --- a cheaper SNN at half the timesteps
  (\texttt{T\ =\ 32}), operating point fit for high VEB \textbf{recall}
  (\textasciitilde0.97 on val). Runs on \textbf{every} beat; emits
  candidate VEBs.
\item
  \textbf{Stage 2 (confirmer)} --- the frozen v1 (\texttt{T\ =\ 64}) at
  a high-\textbf{precision} operating point, run \textbf{only} on
  Stage-1 candidates. A beat is declared VEB iff both stages fire.
\end{itemize}

Stage 1 must be a \emph{different} network from v1 (different \texttt{T}
⇒ different encoding ⇒ partly independent errors); otherwise re-judging
its own candidates adds nothing. Average energy is
\texttt{SynOps(stage1)\ +\ flag\_rate\ ·\ SynOps(stage2\ on\ candidates)},
and because the candidate set is small the heavier confirmer costs
little on average. Both operating points are fit on val only.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Experiments}\label{experiments}

All results use 5 training seeds. The frozen seed is selected by
\textbf{validation} performance; test numbers are reported for that
frozen model and, for stability, as the 5-seed DS2 mean/min. No number
below was used to choose an operating point on the split it is reported
on.

\subsubsection{5.1 Single-stage VEB detection
(val-locked)}\label{single-stage-veb-detection-val-locked}

DS2 VEB detection under each val-fit strategy, 5 seeds, reported as mean
(min):

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Operating point & VEB sensitivity & VEB PPV & SynOps/beat \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{sens-first} & 0.905 (0.878) & 0.539 (0.490) &
\textasciitilde14.3k \\
\textbf{balanced} & 0.857 (0.790) & 0.679 (0.517) &
\textasciitilde14.3k \\
ppv-first & 0.845 (0.771) & 0.700 (0.583) & \textasciitilde14.3k \\
\end{longtable}

The single operating point buys sensitivity \textbf{or} precision, not
both: at sensitivity ≈0.90 the PPV is \textasciitilde0.54, and reaching
PPV \textasciitilde0.68 costs \textasciitilde4 points of sensitivity.
All configurations sit comfortably under the 25k-SynOps budget.

\textbf{Sensitivity--PPV frontier (DS2, oracle sweep).} PPV achievable
at each sensitivity level (range over the 5 seeds; 2 of 5 seeds top out
just below 0.90 sensitivity on DS2, an artifact of the small validation
holdout --- Section 6):

\begin{longtable}[]{@{}lllll@{}}
\toprule\noalign{}
sensitivity ≥ & 0.80 & 0.85 & 0.90 & 0.92 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
VEB PPV & 0.57--0.78 & 0.57--0.69 & 0.56--0.65 & 0.48--0.57 \\
\end{longtable}

Even the oracle frontier does not offer a single point at both ≥0.90
sensitivity and ≥0.60 PPV reliably across seeds --- motivating the
cascade (Section 5.3).

\subsubsection{5.2 Cross-database
generalization}\label{cross-database-generalization}

The frozen NeuroBeat-VEB v1 (sens-first operating point, fit on DS1-val)
applied \textbf{unchanged} to three databases:

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Database & Beats & VEB sensitivity & VEB PPV \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
MIT-BIH DS2 & \textasciitilde50k & 0.894 & 0.490 \\
MIT-BIH SVDB & \textasciitilde184k & 0.892 & 0.361 \\
INCART (12-lead) & \textasciitilde176k & 0.880 & 0.736 \\
\end{longtable}

VEB \textbf{sensitivity holds at 0.88--0.89 across all three databases}
under one frozen operating point --- the detector transfers. PPV varies
with each database's class mix (lower on the supraventricular-dense
SVDB, higher on INCART).

\subsubsection{5.3 Two-stage cascade}\label{two-stage-cascade-1}

A naive cascade --- a T32 screener gating the single frozen v1 ---
reaches DS2 0.883 / 0.622 but at \textbf{19.4k} SynOps, \emph{more} than
single-stage v1, because count-pooling makes the T32 screener as
expensive as the T64 confirmer (Section 6). We therefore use a
\textbf{sparse} screener (2× fewer hidden units, higher spike threshold;
5,987 SynOps/beat) and, since a K-seed ensemble is what actually reaches
the accuracy target, a \textbf{gated 3-seed ensemble confirmer} run only
on the \textasciitilde27\% of beats the screener flags. The ensemble
confirmer requires no extra training --- the three seed models are
already trained for the single-stage study --- and its members' logits
are combined by averaging.

Both operating points are fit on DS1-val only (screener recall target
0.97; confirmer maximizes cascade PPV subject to cascade sensitivity ≥
0.90). The frozen result:

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Database & VEB sensitivity & VEB PPV & flag rate \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{MIT-BIH DS2} & \textbf{0.923} & \textbf{0.616} & 0.271 \\
MIT-BIH SVDB & 0.904 & 0.377 & 0.345 \\
INCART & 0.901 & 0.835 & 0.241 \\
\end{longtable}

Average energy: \texttt{5,987\ +\ 0.271\ ·\ 3\ ·\ 21,433\ =\ 23,385}
SynOps/beat. The cascade \textbf{meets all three targets simultaneously
on DS2} --- sensitivity ≥0.90, PPV ≥0.60, and ≤25k SynOps --- while
holding VEB sensitivity ≥0.90 across all three databases. Ensemble
composition is robust: K∈\{2,3,5\} and different 3-seed subsets all give
DS2 sensitivity ≥0.90 at PPV 0.59--0.64.

\textbf{Accuracy--energy Pareto (DS2 VEB).}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Model & sens / PPV & SynOps/beat \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Single-stage v1 (sens-first) & 0.894 / 0.490 & 14.2k \\
Naive T32 cascade & 0.883 / 0.622 & 19.4k \\
5-seed ensemble (every beat) & 0.932 / 0.595 & \textasciitilde71k \\
\textbf{Gated-ensemble cascade} & \textbf{0.923 / 0.616} &
\textbf{23.4k} \\
\end{longtable}

\subsubsection{5.4 Baselines on the same
split}\label{baselines-on-the-same-split}

We train non-spiking CNN and LSTM baselines under the \textbf{identical}
protocol --- same DS1-train, same beat+RR inputs, same DS1-val
sens-first operating point, same frozen DS2 test, 5 seeds:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Model
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
DS2 VEB sens
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
DS2 VEB PPV
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Ops/beat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Params
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
CNN (1-D, beat+RR) & 0.945 & 0.434 & 356k MACs & 2.9k \\
LSTM (beat+RR) & 0.940 & 0.178 & 4.26M MACs & 17.5k \\
SNN single-stage v1 & 0.894 & 0.490 & 14.2k SynOps & 18k \\
\textbf{SNN gated-ensemble} & \textbf{0.923} & \textbf{0.616} &
\textbf{23.4k SynOps} & --- \\
\end{longtable}

Two observations. First, the dense baselines exhibit the \emph{same}
sensitivity--precision tradeoff --- under the sens-first operating point
they reach high sensitivity (\textasciitilde0.94) but low PPV (0.43 CNN,
0.18 LSTM) --- confirming the tradeoff is a property of single-lead VEB
detection, not of spiking networks. Second, on the operation count that
stands in for energy, the SNN is far cheaper: the CNN and LSTM require
\textasciitilde356k and \textasciitilde4.26M multiply-accumulates per
beat, whereas the SNN performs \textasciitilde14--23k synaptic
operations. A MAC and a SynOp are not identical units --- a SynOp is a
single accumulate triggered by a spike, and on neuromorphic hardware is
cheaper than a clocked MAC --- so the \textasciitilde15--180×
operation-count gap is a conservative lower bound on the energy
advantage. The gated-ensemble SNN also attains the best precision
(0.616) of any model at comparable sensitivity.

\subsubsection{5.5 Supraventricular beats: a single-lead
limitation}\label{supraventricular-beats-a-single-lead-limitation}

SVEB detection is hard on one lead: supraventricular ectopic beats often
differ from normal beats mainly in timing, not morphology, and DS1
contains few of them. As an incidental output of the VEB model, DS2 SVEB
sensitivity is \textasciitilde0.13; adding supraventricular-rich svdb to
training raises it to \textasciitilde0.24, and the T96 lever to
\textasciitilde0.27 --- still far from usable.

We then trained a \textbf{dedicated SVEB specialist} (T96, svdb + INCART
augmentation, SVEB-only operating point, 5 seeds). It does \textbf{not}
yield a usable detector:

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Seed regime & DS2 SVEB sens & DS2 SVEB PPV & DS2 VEB sens \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
calibrated (seeds 0--2) & 0.16--0.27 & 0.07--0.15 & collapses \\
degenerate (seeds 3--4) & 0.82--0.94 & 0.04--0.06 & 0.18--0.24 \\
\end{longtable}

The high-sensitivity seeds reach it only by flagging almost everything
as SVEB (PPV ≤0.06) and in doing so cannibalize the ventricular class.
DS2 SVEB PPV never exceeds \textasciitilde0.15 at any operating point,
and the result is wildly seed-unstable. The same architecture reached
SVEB sensitivity 0.62 on 12-lead INCART, so the limit is \textbf{lead
count and data, not capacity}. Multi-lead input (MIT-BIH provides two
leads; supraventricular beats separate better on some leads) is the
natural next lever and is left to future work. This is a genuine
negative result: on a single lead, supraventricular detection at usable
precision is out of reach here, which is precisely why it must not be
allowed to compromise the VEB operating point.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Discussion}\label{discussion}

\textbf{Count-pooling makes energy timestep-independent --- with two
consequences.} Under count-pooled delta encoding the total number of
input spikes is approximately conserved across timesteps \texttt{T}; the
dominant SynOps term (input-spikes × hidden) therefore barely changes
with \texttt{T}. (i) \emph{Time resolution is a free discrimination
lever}: raising \texttt{T} from 32 to 64 sharpens separation of ectopic
morphologies at almost no energy cost. (ii) \emph{You cannot build a
cheap screener by lowering \texttt{T}} --- a T32 network costs
essentially the same as a T64 one. A genuinely sparse screener must
instead reduce input channels, hidden units, or raise the spike
threshold. This is why our two-stage screener is sparse-by-construction
rather than merely shorter.

\textbf{The shortfall of a single model is calibration variance, not
capacity.} The DS2 oracle frontier shows ≥0.90 sensitivity at ≥0.60 PPV
is achievable, yet a single model calibrated on the small validation
holdout rarely lands there and varies widely across seeds. A K-seed deep
ensemble both averages away model variance and calibrates cleanly,
reaching the target corner (0.918 / 0.634 on DS2 for a 3-seed ensemble,
with \textasciitilde0.90 sensitivity across all three databases). The
two-stage \textbf{gated-ensemble} cascade delivers this accuracy within
the energy budget by running the ensemble only on flagged candidates.

\textbf{The honest protocol costs headline accuracy but buys trust.}
Freezing the operating point on val rather than test lowers reported
numbers relative to test-tuned protocols, but the cross-database results
show the frozen point transfers, which test-tuned numbers cannot
promise.

\textbf{Limitations.} Single-lead SVEB detection is inherently hard; our
energy proxy is SynOps, not measured hardware joules; and the small
validation holdout makes single-model operating points noisy (mitigated
by the ensemble). A larger or cross-validated DS1 holdout is likely to
tighten single-model calibration.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. Conclusion}\label{conclusion}

NeuroBeat is a compact, energy-accounted spiking VEB detector evaluated
under a conservative, leak-free inter-patient protocol with
cross-database external validation. A two-stage cascade delivers high
sensitivity and precision within a 25k-SynOps/beat budget. We release
code, frozen weights, and a live dashboard, and we treat
supraventricular detection as a separate, data-limited problem.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{8. Reproducibility}\label{reproducibility}

\begin{itemize}
\tightlist
\item
  Repository: github.com/Talch87/neuro-beat (AGPL-3.0 / commercial dual
  license)
\item
  Live dashboard: talch87.github.io/neuro-beat
\item
  Frozen artifact:
  \texttt{models/neurobeat-veb-v1/\{weights.pt,\ operating\_point.json\}}
\item
  Protocol harness: \texttt{experiments/freeze\_veb\_v1.py},
  \texttt{experiments/gated\_ensemble\_veb.py},
  \texttt{experiments/baselines\_veb.py}
\item
  All experiments: 5 seeds, operating point fit on DS1-val only.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{References}\label{references}

\emph{Compiled and cross-checked via literature search (July 2026).
Author lists for the SNN-ECG works marked † should be confirmed against
the source before submission.}

\begin{itemize}
\tightlist
\item
  \textbf{{[}AAMI-EC57{]}} ANSI/AAMI EC57. \emph{Testing and Reporting
  Performance Results of Cardiac Rhythm and ST-Segment Measurement
  Algorithms.} Association for the Advancement of Medical
  Instrumentation.
\item
  \textbf{{[}deChazal2004{]}} P. de Chazal, M. O'Dwyer, R. B. Reilly.
  ``Automatic classification of heartbeats using ECG morphology and
  heartbeat interval features.'' \emph{IEEE Trans. Biomed. Eng.},
  51(7):1196--1206, 2004. doi:10.1109/TBME.2004.827359.
\item
  \textbf{{[}Silva2025{]}} G. Silva, P. Silva, G. Moreira, V. Freitas,
  J. Gertrudes, E. Luz. ``A Systematic Review of ECG Arrhythmia
  Classification: Adherence to Standards, Fair Evaluation, and Embedded
  Feasibility.'' arXiv:2503.07276, 2025.
\item
  \textbf{{[}Moody2001{]}} G. B. Moody, R. G. Mark. ``The impact of the
  MIT-BIH Arrhythmia Database.'' \emph{IEEE Eng. Med. Biol. Mag.},
  20(3):45--50, 2001.
\item
  \textbf{{[}Goldberger2000{]}} A. L. Goldberger et al.~``PhysioBank,
  PhysioToolkit, and PhysioNet.'' \emph{Circulation},
  101(23):e215--e220, 2000.
\item
  \textbf{{[}Neftci2019{]}} E. O. Neftci, H. Mostafa, F. Zenke.
  ``Surrogate Gradient Learning in Spiking Neural Networks.'' \emph{IEEE
  Signal Process. Mag.}, 36(6):51--63, 2019.
\item
  \textbf{{[}Eshraghian2023{]}} J. K. Eshraghian et al.~``Training
  Spiking Neural Networks Using Lessons From Deep Learning.''
  \emph{Proc. IEEE}, 111(9):1016--1054, 2023.
\item
  \textbf{{[}Davies2018{]}} M. Davies et al.~``Loihi: A Neuromorphic
  Manycore Processor with On-Chip Learning.'' \emph{IEEE Micro},
  38(1):82--99, 2018.
\item
  \textbf{{[}SNN-ECG-BSPC2021{]}} † ``Energy efficient ECG
  classification with spiking neural network.'' \emph{Biomed. Signal
  Process. Control}, 2021. (ScienceDirect S1746809420303098.)
\item
  \textbf{{[}Chu2022{]}} † ``A Neuromorphic Processing System With
  Spike-Driven SNN Processor for Wearable ECG Classification.''
  \emph{IEEE Trans. Biomed. Circuits Syst.}, 2022. (PubMed 35802543.)
\item
  \textbf{{[}SparrowSNN2024{]}} † ``SparrowSNN: A Hardware/Software
  Co-design for Energy Efficient ECG Classification.'' arXiv:2406.06543,
  2024.
\item
  \textbf{{[}AxonalDelays2025{]}} † ``Robust ECG signal classification
  using spiking neural networks with axonal delays.''
  \emph{Neurocomputing}, 2025. (ScienceDirect S0925231225029315.)
\end{itemize}

\end{document}
